---
title: "Astrangia poculata distribution"
author: "jillashey"
date: "2023-05-11"
output: html_document
---

This script creates a map of Astrangia distribution based on data collected from OBIS and xxxxx. 

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages(c("cowplot", "googleway", "ggplot2", "ggrepel", "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))

library("cowplot")
library("googleway")
library("ggplot2")
library("ggrepel")
library("ggspatial")
#library("libwgeom")
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("tidyverse")
```

## Load and subset data 
```{r}
## data from OBIS
obis <- read.csv("data/OBIS/Occurrence.csv")
obis <- obis %>%
  select(id, decimallongitude, decimallatitude, date_year, scientificname, originalscientificname, institutioncode, sst, basisofrecord, occurrencestatus, waterbody, country, stateprovince) %>% # select specific columns of interest
  drop_na(decimallongitude) %>% # remove data that does not have longitude values
  drop_na(decimallatitude) # remove data that does not have latitude values

## data from GBIF
gbif <- read.delim("data/GBIF/occurrence.txt")
gbif <- gbif %>%
  select(gbifID, rightsHolder, basisOfRecord, occurrenceStatus, year, continent, countryCode, stateProvince, decimalLongitude, decimalLatitude, scientificName, acceptedScientificName, collectionCode, datasetName) %>% # select specific columns of interest
  drop_na(decimalLongitude) %>% # remove data that does not have longitude values
  drop_na(decimalLatitude) # remove data that does not have latitude values
```

## Examine where the data came from & how the data was collected 
```{r}
## OBIS
unique(obis$institutioncode)
#  [1] "WoodsHoleBioSurvey"                             "HRI"                                            ""                                              
#  [4] "Smithsonian Environmental Research Center"      "SMCC"                                           "TCWC"                                          
#  [7] "USNM"                                           "NOAA, NMFS, Northeast Fisheries Science Center" "EMAP_NCA"                                      
# [10] "HEX"                                            "OGL"                                            "UF"    

unique(obis$basisofrecord)
# [1] "HumanObservation"  "PreservedSpecimen"


## GBIF
unique(gbif$datasetName)
# [1] "iNaturalist research-grade observations"     ""                                            "NMNH Extant Biology"                        
# [4] "NMNH Material Samples (USNM)"                "Ocean Genome Legacy Collection"              "NCSM Non-molluscan Invertebrates Collection"

unique(gbif$basisOfRecord)
# [1] "HUMAN_OBSERVATION"  "PRESERVED_SPECIMEN" "FOSSIL_SPECIMEN"    "MATERIAL_SAMPLE"   
```

I am only interested in mapping the points that came from 'human observations'. Preserved samples or fossil specimans might skew the data depending on where they were collected. 

## Filter by Human Observation only 
```{r}
obis_human <- obis %>%
  filter(basisofrecord == "HumanObservation")
unique(obis_human$institutioncode)

gbif_human <- gbif %>%
  filter(basisOfRecord == "HUMAN_OBSERVATION")
unique(gbif_human$datasetName)
```

## Combine datasets 
```{r}
obis_human$Source <- "OBIS"
obis_human <- obis_human %>%
  select(decimallongitude, decimallatitude, date_year, basisofrecord, Source)

gbif_human$Source <- "GBIF"
gbif_human <- gbif_human %>%
  select(basisOfRecord, year, decimalLongitude, decimalLatitude, Source) %>%
  relocate(decimalLongitude, decimalLatitude, year, basisOfRecord, Source) %>%
  rename("decimallongitude" = "decimalLongitude", "decimallatitude" = "decimalLatitude", "date_year" = "year", "basisofrecord" = "basisOfRecord")
  
all <- rbind(obis_human, gbif_human)
```

## Plot individual sources 
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")

## OBIS
obis_plot <- ggplot(data = world) +
  geom_sf() +
  geom_point(data = obis_human, aes(x = decimallongitude, y = decimallatitude, shape = Source, color = Source), size = 2, color = "black") +
  coord_sf(xlim = c(-100, -55), ylim = c(13, 45), expand = FALSE) +
  ylab("Latitude") +
  xlab("Longitude") +
  theme_linedraw() +
  theme(legend.position = "none")

obis_plot + annotate(geom = "text", x = -90, y = 37, label = "North America") +
  annotate(geom = "text", x = -65, y = 28, label = "Atlantic Ocean")

## GBIF
gbif_plot <- ggplot(data = world) +
  geom_sf() +
  geom_point(data = gbif_human, aes(x = decimallongitude, y = decimallatitude, shape = Source, color = Source), size = 2, color = "black") +
  coord_sf(xlim = c(-100, -55), ylim = c(13, 45), expand = FALSE) +
  ylab("Latitude") +
  xlab("Longitude") +
  #geom_text(x = -90, y = 37, label = "test") +
  theme_linedraw() +
  theme(legend.position = "none")

gbif_plot + annotate(geom = "text", x = -90, y = 37, label = "North America") +
  annotate(geom = "text", x = -65, y = 28, label = "Atlantic Ocean")
```

## Plot both datasets 
```{r}
both <- ggplot(data = world) +
  geom_sf() +
  geom_point(data = all, aes(x = decimallongitude, y = decimallatitude, shape = Source, color = Source), size = 2) +
  coord_sf(xlim = c(-100, -60), ylim = c(15, 45), expand = FALSE) +
  ylab("Latitude") +
  xlab("Longitude") +
  theme_linedraw()
both
```

Weird that the GBIF dataset has less data points but in the plot, it looks like it has more data points compared to OBIS. Maybe the OBIS dataset has a lot of duplicates? 
```{r}
length(unique(obis_human$decimallongitude)) # 152 unique longitude measurements 
length(unique(obis_human$decimallatitude)) # 139 unique latitude measurements 

length(unique(gbif_human$decimallongitude)) # 393 unique longitude measurements 
length(unique(gbif_human$decimallatitude)) # 395 unique latitude measurements 

## Okay so the GBIS dataset has more unique lat/long measurements 
```

Also to note - GBIF has much more recent data. The most recent GBIF observation is from 2023, whereas the most recent OBIS observation is from 1993. OBIS seems like it has more historical data from the early 1900s. 


Next step - make plots that include all samples, not just those that are from Human Observations 

## Plot with all data 
```{r}
## OBIS
ggplot(data = world) +
  geom_sf() +
  geom_point(data = obis, aes(x = decimallongitude, y = decimallatitude, color = basisofrecord), size = 2) +
  #coord_sf(xlim = c(-100, -50), ylim = c(-15, 45), expand = FALSE) +
  ylab("Latitude") +
  xlab("Longitude") +
  theme_linedraw() 

## GBIF
ggplot(data = world) +
  geom_sf() +
  geom_point(data = gbif, aes(x = decimalLongitude, y = decimalLatitude, color = basisOfRecord), size = 2) +
  #coord_sf(xlim = c(-100, 0), ylim = c(-15, 45), expand = FALSE) +
  ylab("Latitude") +
  xlab("Longitude") +
  theme_linedraw()
```

For the human observation in Portugal, this was the notes: "tentative ID. as far as I know, this species has not been recorded from Europe, but it would be a species likely to be introduced via shipping, etc. the corallite and colony shape all seem appropriate, and it does commonly wash ashore. https://www.inaturalist.org/observations/59814300 https://www.inaturalist.org/observations/29579403". It was done on iNaturalist (most of GBIF info comes from there). 


### Notes 

floor function for year - group by decade 
is there any overlap in data points between databases?
Kernel density estimate - hot spots of where species are reported 

## Group and plot by decade 
```{r}
test <- gbif_human %>%
  mutate(decade = floor(date_year/10)*10) %>%
  group_by(decade)

## GBIF
test_plot <- ggplot(data = world) +
  geom_sf() +
  geom_point(data = test, aes(x = decimallongitude, y = decimallatitude, color = decade), size = 2) +
  coord_sf(xlim = c(-100, -55), ylim = c(13, 45), expand = FALSE) +
  ylab("Latitude") +
  xlab("Longitude") +
  #geom_text(x = -90, y = 37, label = "test") +
  theme_linedraw() +
  theme(legend.position = "none")

test_plot + annotate(geom = "text", x = -90, y = 37, label = "North America") +
  annotate(geom = "text", x = -65, y = 28, label = "Atlantic Ocean")

```



Merge datasets 
```{r}
obis_human$coordinates <- paste0(obis_human$decimallatitude, sep = ", ", obis_human$decimallongitude)
length(unique(obis_human$coordinates))

gbif_human$coordinates <- paste0(gbif_human$decimalLatitude, sep = ", ", gbif_human$decimalLongitude)
length(unique(gbif_human$coordinates))

merge <- full_join(obis_human, gbif_human, by = "coordinates") %>%
  drop_na(decimalLatitude, decimallatitude, decimalLongitude, decimallongitude)

length(unique(merge$decimallongitude))
length(unique(merge$decimallatitude))

merge <- unique(merge)
```

# Plot merge datasets 
```{r}
test_plot <- ggplot(data = world) +
  geom_sf() +
  geom_point(data = merge, aes(x = decimallongitude, y = decimallatitude), size = 2) +
  coord_sf(xlim = c(-90, -70), ylim = c(20, 45), expand = FALSE) +
  ylab("Latitude") +
  xlab("Longitude") +
  #geom_text(x = -90, y = 37, label = "test") +
  theme_linedraw() +
  theme(legend.position = "none")

test_plot + annotate(geom = "text", x = -90, y = 37, label = "North America") +
  annotate(geom = "text", x = -65, y = 28, label = "Atlantic Ocean")
```






