---
title: "Literature data"
author: "jillashey"
date: "2023-05-10"
output: html_document
---

This script will plot the cummulative distribution of the peer-reviewed literature on **Astrangia poculata** prior to December 2023. This data does not include dissertations, theses, etc. 

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggpubr)
library(cowplot)
library(gridExtra)
library(plotly)
library(ggrepel)
```

## Load lit data by year 
```{r}
lit <- read.csv("../data/LiteratureByYear.csv")

# Remove NAs
lit <- na.omit(lit)
```

## Plot lit data by year 
```{r}
lit_by_year <- ggplot(lit, aes(y = PublicationNo., x = Year)) +
  geom_bar(stat = "identity", position = "dodge", fill = "gray", color = "black", width = 1, size = 2) +
  theme_minimal() +
  ylab("Number of Publications\n") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, 6), breaks = seq(0,6,1)) +
  scale_x_continuous(breaks = seq(1925,2023,5)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 65,
                                   #vjust = 0.5,
                                   hjust= 1,
                                   size=45,
                                   colour = "black"),
        axis.title.x = element_text(size = 0),
        axis.text.y = element_text(size = 50, colour = "black"),
        axis.title.y = element_text(size = 45, face = "bold"),
        axis.line = element_line(colour = 'black', size = 1),
        axis.ticks.x = element_line(size = 1, color = "black"),
        axis.ticks.length = unit(0.5, "cm"),
        #plot.title = element_text(size = 100, hjust = 0.5),
        plot.margin = margin(3,3,3,3, "cm")); lit_by_year

# Save plot 
ggsave("../output/Literature_By_Year.pdf", lit_by_year, width = 40, height = 15)
ggsave("../output/Literature_By_Year.png", lit_by_year, width = 40, height = 15)
```



## Load lit data by focal area
```{r}
lit_focal <- read.csv("../data/LiteratureByFocalArea.csv")

# Replace "Habitat" with "Habitat & Ecology"
lit_focal$Primary.Focal.Area <- gsub("Habitat", "Habitat & Ecology", lit_focal$Primary.Focal.Area)
```

## Calculate percentages of each focal area 
```{r}
lit_focal.percent <- lit_focal %>%
  group_by(Primary.Focal.Area) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```

## Plot lit data by focal area 
```{r}
# Get the positions
df2 <- lit_focal.percent %>% 
  mutate(csum = rev(cumsum(rev(count))), 
         pos = count/2 + lead(csum, 1),
         pos = if_else(is.na(pos), count/2, pos))

# Edit positions for certain focal areas, based on looking at the plot 
df2$pos <- c(54, 49, 33, 18, 14.5, 6.5)

lit_by_focal_area <- ggplot(lit_focal.percent, aes(x = "" , y = count, fill = fct_inorder(Primary.Focal.Area))) +
  geom_col(width = 1.5, color = "black", linewidth = 1.5) +
  #coord_polar(theta = "y") +
  coord_polar("y", start=0) + 
  scale_fill_brewer(palette = "Dark2") +
  geom_label_repel(data = df2,
                   aes(y = pos, label = paste(Primary.Focal.Area,"(", round(percentage, 1), "%)")),
                   size = 15, nudge_x = 1, show.legend = FALSE,
                   box.padding = 0.5, box.color = "black", box.alpha = 0.5, 
                   label.r = unit(1, "lines"),  # Adjust the radius as needed
                   label.padding = unit(1,"lines"), segment.color = "black") +
  theme_void() + 
  theme(legend.position = "none"); lit_by_focal_area

# Save plot 
ggsave("../output/Literature_By_FocalArea.pdf", lit_by_focal_area, width = 20, height = 15, units = c("in"))
ggsave("../output/Literature_By_FocalArea.png", lit_by_focal_area, width = 20, height = 15, units = c("in"))
```

## Calculate percentages of each location
```{r}
# Select location only and calculate percentages
lit_focal_location.percent <- lit_focal %>%
  na.omit(Location) %>%
  select(Location) %>%
  mutate(Location = trimws(Location)) %>%
  separate_rows(Location, sep = "; ") %>%
  group_by(Location) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```

## Plot lit data by location 


STILL NEED TO WORK ON THIS PLOT 
```{r}
# Get the positions
df3 <- lit_focal_location.percent %>% 
  mutate(csum = rev(cumsum(rev(count))), 
         pos = count/2 + lead(csum, 1),
         pos = if_else(is.na(pos), count/2, pos))

# Edit positions for certain focal areas, based on looking at the plot 
df3$pos <- c(59.5, # Delaware 
             59, # Florida  
             50, # Georgia 
             53, # Maryland 
             46.5, # Massachusetts
             40, # New York 
             36, # North Carolina 
             19, # Rhode Island 
             1.5) # Virginia

lit_by_location <- ggplot(lit_focal_location.percent, aes(x = "" , y = count, fill = fct_inorder(Location))) +
  geom_col(width = 1.5, color = "black", linewidth = 1.5) +
  #coord_polar(theta = "y") +
  coord_polar("y", start=0) + 
  scale_fill_manual(values = c("#7f7f7f", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2", "#1f77b4", "#bcbd22")) +
  geom_label_repel(data = df3,
                   aes(y = pos, label = paste(Location,"(", round(percentage, 1), "%)")),
                   size = 15, nudge_x = 1, show.legend = FALSE,
                   box.padding = 0.5, box.color = "black", box.alpha = 0.5, 
                   label.r = unit(1, "lines"),  # Adjust the radius as needed
                   label.padding = unit(1,"lines"), segment.color = "black") +
  theme_void() + 
  theme(legend.position = "none"); lit_by_location

# Save plot 
ggsave("../output/Literature_By_Location.pdf", lit_by_location, width = 20, height = 15, units = c("in"))
ggsave("../output/Literature_By_Location.png", lit_by_location, width = 20, height = 15, units = c("in"))
```

## Arrange plots into a single figure

NEED TO WORK ON THIS









